---
title: "Teste Melhor Impressão de Barcharts"
author: "José Elvano Moraes"
date: "4/15/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo=FALSE}
suppressPackageStartupMessages(expr = library(tidyverse))
suppressPackageStartupMessages(expr = library(bnlearn))
suppressPackageStartupMessages(expr = library(bnstruct))
suppressPackageStartupMessages(expr = library(readr))
suppressPackageStartupMessages(expr = library(Rgraphviz))
suppressPackageStartupMessages(expr = library(gRain))
suppressPackageStartupMessages(expr = library(Rmpfr))
```


```{r echo=FALSE}
# library(readr)
#  dd <- read_delim("dados/INFLUD21-05-04-2021 copy.csv",
#     ";", escape_double = FALSE, trim_ws = TRUE)
```


```{r echo=FALSE}
# colunas <- c(
#  'NU_IDADE_N',
#  'TP_IDADE',
#  'FEBRE',
#  'TOSSE',
#  'GARGANTA',
#  'DISPNEIA',
#  'DESC_RESP',
#  'SATURACAO',
#  'DIARREIA',
#  #'VOMITO',
#  'OUTRO_SIN',
#  #'HOSPITAL',
#  'EVOLUCAO',
#  'RENAL',
#  'DIABETES',
#  'OBESIDADE',
#  'PERD_OLFT',
#  'PERD_PALA',
#  #'VACINA',
#  'CLASSI_FIN',
#  #'CS_SEXO',
#  #'CS_GESTANT',
#  #'PUERPERA' ,
#  'CARDIOPATI',
#  'SUPORT_VEN',
#  'HEMATOLOGI', 
#  #'SIND_DOWN' ,
#  #'HEPATICA' ,
#  'ASMA'#,
#  #'ANTIVIRAL'
#  #'TP_ANTIVIR'
#  #'RAIOX_RES'
# )
```

```{r echo=FALSE}
#separar grupo de variaveis para analise
# ddf <- dd[, colunas] %>%
#   rename(IDADE = NU_IDADE_N) %>%
#   filter(CLASSI_FIN == 5) %>%
#   na.omit(ddf) %>%
#   filter(TP_IDADE == 3) %>%
#   mutate_if(is.double,as.factor)  %>%
#   mutate_if(is.character,as.factor) %>%
#   select(-TP_IDADE)
# 
#   ddf$CLASSI_FIN <- NULL
# 
#   ddf$IDADE = as.double(ddf$IDADE)
#   ddf <- as.data.frame(ddf)
# rm(dd)
# glimpse(ddf)
#saveRDS(ddf, 'ddf.rds')
  
  ddf <- readRDS('ddf.rds')
```

```{r echo=FALSE}
# discretireadR ze idade

ddf <- discretize(ddf, method = 'interval') 
glimpse(ddf)
```


```{r echo=FALSE}
s1 <- sample_frac(ddf, .8, FALSE)
#s2 <- sample_frac(df, .9, FALSE)
#s3 <- sample_frac(df, .9, FALSE)
#s4 <- sample_frac(df, .9, FALSE)

s1$HEMATOLOGI = NULL

wl = matrix(c("OBESIDADE", "DIABETES",
              "IDADE", "DIABETES",
              "IDADE", "CARDIOPATI",
              "IDADE", "SUPORT_VEN",
              "DIABETES", "RENAL",
              "DIABETES", "CARDIOPATI",
              "RENAL", "CARDIOPATI"),
        ncol = 2, byrow = TRUE, dimnames = list(NULL, c("from", "to")))
cat("White list:")
wl

bn1 <- mmhc(s1, whitelist = wl)
#bn2 <- mmhc(s2)
#bn3 <- mmhc(s3)
#bn4 <- mmhc(s4)
```


```{r echo=FALSE}
#par(mfrow=c(2,2))
graphviz.plot(bn1, shape='rectangle', main = '...')
# graphviz.plot(bn2, shape='rectangle')
# graphviz.plot(bn3, shape='rectangle')
# graphviz.plot(bn4, shape='rectangle')

```


```{r echo=FALSE}
fitted.1 <- bn.fit(bn1, s1)
# fitted.2 <- bn.fit(bn2, s2)
# fitted.3 <- bn.fit(bn3, s3)
# fitted.4 <- bn.fit(bn4, s4)

#par(mfrow=c(2,2))
graphviz.chart(fitted.1, type = "barprob", col = "darkblue", bg = "azure", bar.col = "darkblue",
               main = "Rede de probabilidasdes")
# graphviz.chart(fitted.2, type = "barprob", col = "darkblue", bg = "azure", bar.col = "darkblue")
# graphviz.chart(fitted.3, type = "barprob", col = "darkblue", bg = "azure", bar.col = "darkblue")
# graphviz.chart(fitted.4, type = "barprob", col = "darkblue", bg = "azure", bar.col = "darkblue")

```

---------

## Bootstraping
```{r}
boots.trap  <- 100
str.diff = suppressMessages(boot.strength(s1, R = boots.trap, algorithm = "mmhc"))
cat(paste('Threshold: ', attr(str.diff, "threshold")))
avg.diff = averaged.network(str.diff)
strength.plot(avg.diff, str.diff, shape = "rectangle", main = paste("Iterações = ", boots.trap))
```


How can we compare the averaged network (avg.diff) with the network we originally learned in from all the data? The most qualitative way is to plot the two networks side by side, with the nodes in the same positions, and highlight the arcs that appear in one network and not in the other, or that appear with different directions.

```{r}
par(mfrow = c(1, 2))
graphviz.compare(avg.diff, bn1, shape = "rectangle", main = c("DAG médio", "DAG único"))
```

It is also a good idea to look at the threshold with respect to the distribution of the arc strengths
```{r}
plot(str.diff)
abline(v = 0.75, col = "tomato", lty = 2, lwd = 2)
abline(v = 0.85, col = "steelblue", lty = 2, lwd = 2)
```

The simpler network we obtain by setting threshold = 0.8 in averaged.network() is shown below; it is certainly easier to reason with from a qualitative point of view.
```{r}
par(mfrow = c(1, 2))
avg.simpler = averaged.network(str.diff, threshold = 0.75)
strength.plot(avg.diff, str.diff, shape = "rectangle", main = paste("Iterações = ", boots.trap, " Thr = ", attr(str.diff, "threshold")))
strength.plot(avg.simpler, str.diff, shape = "rectangle", main = 'Iterações = 100 Thr = 0.75')


```

